---
title: "Final Assignment"
author: "Venkat Ganesh"
date: "1/4/2021"
output: html_document
---
## Summary 

This problem involved predicting the type of activity based on predictor variables obtained from measurements of accelerometers. The first step of my approach was to see if I could do any data cleaning. It was immediately apparent that there were columns with NA measurements, so I wanted to exclude these from my predictors. I also observed some variables which I did not believe would have any predictive capabilities so I removed those. Lastly I used the nearZeroVar function to remove predictors with variances near 0, as these would not have much predictive power. After cleaning, my first modeling approach was to fit a classification tree. This would give us a relatively fast model fit, along with outputting interpretable results. The model ended up being not very accurate, getting an accuracy around 50%. So I instead tried a Random Forest model, which ended up giving me a 99.5% accuracy on the holdout set, and yielded 100% results in the out of sample test set provided for us.

## Read Data and load packages
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(dplyr)
library(rattle)
train_raw <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
test_raw <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

```

## data cleaning

```{r}
head_data <- head(train_raw,5)
train <- train_raw[, colSums(is.na(train_raw)) == 0] 
test <- test_raw[, colSums(is.na(test_raw)) == 0] 
train <- train[,!(colnames(train) %in% c("X","user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window","num_window"))]
train$classe <- as.factor(train$classe)
nearzerocolumns <- nearZeroVar(train)
train <- train[, -nearzerocolumns]

```

## Classification Tree Modeling

```{r}
set.seed(14231)
trainIndex <- createDataPartition(train$classe, p = 0.75)[[1]]
validation <- train[-trainIndex,]
train <- train[trainIndex,]
modelFit <- train(classe~.,method = "rpart",data = train)
modelTest <- predict(modelFit, validation)
confusionMatrix(modelTest, validation$classe)
```

## Classification tree

```{r}
fancyRpartPlot(modelFit$finalModel)
```

## Random Forest Modeling

```{r}
set.seed(14232)
modelFit_rf <- train(classe~.,method = "rf",data = train, allowParallel=TRUE)
modelTest_rf <- predict(modelFit_rf, validation)
confusionMatrix(modelTest_rf, validation$classe)
```


# Out of Sample Results
```{r}
predict(modelFit_rf, test)
```


